\documentclass[a5paper,12pt]{article}
\usepackage[bg-none]{lib/rpg-book} % Options: bg-a4, bg-letter, bg-full, bg-print, bg-none.
\usepackage[polish]{babel}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{tabularx}

\usepackage{fouriernc}
\usepackage{parskip}
\usepackage{booktabs}
\usepackage{array}

\usepackage{pgfpages}
   \pgfpagesuselayout{2 on 1}[a4paper,landscape,border shrink=5mm]

\title{Trzy Rudzielce i Elf}
\date{\today}
\author{The Dungeon Mistress, Vivien, Yeleda, Calion, Garret}

% Start document
\begin{document}
\fontfamily{ppl}\selectfont % Set text font

% Your content goes here

\pagenumbering{gobble}

\section*{Message}

\begin{table}[h]
   \centering
   \begin{tabular}{|l|r|}
      \hline
      Level        & Cantrip \\
      \hline
      School       & Transmutation \\
      \hline
      Casting Time & Action \\
      \hline
      Range        & 120 ft \\
      \hline
      Components   & V, S, M \\
      \hline
      Duration     & 1 round\\
      \hline
      Area         & Target \\
      \hline
   \end{tabular}
\end{table}

\paragraph{}
You point your finger toward a creature within range and whisper a message. The target (and only the target) hears the message and can reply in a whisper that only you can hear.

\paragraph{}
You can cast this spell through solid objects if you are familiar with the target and know it is beyond the barrier. Magical silence, 1 foot of stone, 1 inch of common metal, a thin sheet of lead, or 3 feet of wood blocks the spell. The spell doesn't have to follow a straight line and can travel freely around corners or through openings.

\newpage


\section*{Minor Illusion}
\begin{table}[h]
   \centering
   \begin{tabular}{|l|r|}
      \hline
      Level        & Cantrip \\
      \hline
      School       & Illusion\\
      \hline
      Casting Time & Action \\
      \hline
      Range        & 30 ft \\
      \hline
      Components   & S, M \\
      \hline
      Duration     & 1 minute\\
      \hline
      Area         & Target \\
      \hline
   \end{tabular}
\end{table}

\paragraph{}
You create a sound or an image of an object within range that lasts for the duration. The illusion also ends if you dismiss it as an action or cast this spell again.

\paragraph{}
If you create a sound, its volume can range from a whisper to a scream. It can be your voice, someone else's voice, a lion's roar, a beating of drums, or any other sound you choose. The sound continues unabated throughout the duration, or you can make discrete sounds at different times before the spell ends.

\paragraph{}
If you create an image of an object--such as a chair, muddy footprints, or a small chest--it must be no larger than a 5-foot cube. The image can't create sound, light, smell, or any other sensory effect. Physical interaction with the image reveals it to be an illusion, because things can pass through it.

\paragraph{}
If a creature uses its action to examine the sound or image, the creature can determine that it is an illusion with a successful Intelligence (Investigation) check against your spell save DC. If a creature discerns the illusion for what it is, the illusion becomes faint to the creature.

\newpage

\section*{Vicious Mockery}
\begin{table}[h]
   \centering
   \begin{tabular}{|l|r|}
      \hline
      Level        & Cantrip \\
      \hline
      School       & Enchantment\\
      \hline
      Casting Time & Action \\
      \hline
      Range        & 60 ft \\
      \hline
      Components   & V \\
      \hline
      Duration     & Instantaneous\\
      \hline
      Area         & Target \\
      \hline
   \end{tabular}
\end{table}

\paragraph{}
You unleash a string of insults laced with subtle enchantments at a creature you can see within range. If the target can hear you (though it need not understand you), it must succeed on a Wisdom saving throw or take 1d4 psychic damage and have disadvantage on the next attack roll it makes before the end of its next turn.

\paragraph{}
This spell's damage increases by 1d4 when you reach 5th level (2d4), 11th level (3d4), and 17th level (4d4).

\newpage

\section*{Dissonant Whispers}
\begin{table}[h]
   \centering
   \begin{tabular}{|l|r|}
      \hline
      Level        & I \\
      \hline
      School       & Enchantment\\
      \hline
      Casting Time & Action \\
      \hline
      Range        & 60 ft \\
      \hline
      Components   & V \\
      \hline
      Duration     & Instantaneous\\
      \hline
      Area         & Target \\
      \hline
   \end{tabular}
\end{table}

\paragraph{}
You whisper a discordant melody that only one creature of your choice within range can hear, wracking it with terrible pain. The target must make a wisdom saving throw. On a failed save, it takes 3d6 psychic damage and must immediately use its reaction, if available, to move as far as its speed allows away from you. The creature doesn’t move into obviously dangerous ground, such as a fire or a pit. On a successful save, the target takes half as much damage and doesn’t have to move away. A deafened creature automatically succeeds on the save.

\paragraph{At higher levels:}
When you cast this spell using a spell slot of 2nd level or higher, the damage increases by 1d6 for each slot level above 1st.

\newpage

\section*{Faerie Fire}
\begin{table}[h]
   \centering
   \begin{tabular}{|l|r|}
      \hline
      Level        & I \\
      \hline
      School       & Evocation \\
      \hline
      Casting Time & Action \\
      \hline
      Range        & 60 ft \\
      \hline
      Components   & V \\
      \hline
      Duration     & Concentration, up to 1 minute\\
      \hline
      Area         & Cube, 20 ft \\
      \hline
   \end{tabular}
\end{table}

\paragraph{}
Each object in a 20-foot cube within range is outlined in blue, green, or violet light (your choice). Any creature in the area when the spell is cast is also outlined in light if it fails a Dexterity saving throw. For the duration, objects and affected creatures shed dim light in a 10-foot radius.

\paragraph{}
Any attack roll against an affected creature or object has advantage if the attacker can see it, and the affected creature or object can't benefit from being invisible.

\newpage

\section*{Healing Word}
\begin{table}[h]
   \centering
   \begin{tabular}{|l|r|}
      \hline
      Level        & I \\
      \hline
      School       & Evocation \\
      \hline
      Casting Time & Bonus Action \\
      \hline
      Range        & 60 ft \\
      \hline
      Components   & V \\
      \hline
      Duration     & Instantaneous \\
      \hline
      Area         & Target \\
      \hline
   \end{tabular}
\end{table}

\paragraph{}
A creature of your choice that you can see within range regains hit points equal to 1d4 + your spellcasting ability modifier. This spell has no effect on undead or constructs.

\paragraph{At higher levels:}
When you cast this spell using a spell slot of 2nd level or higher, the healing increases by 1d4 for each slot level above 1st.

\newpage

\section*{Tasha's Hideous Laughter}
\begin{table}[h]
   \centering
   \begin{tabular}{|l|r|}
      \hline
      Level        & I \\
      \hline
      School       & Enchantment \\
      \hline
      Casting Time & Action \\
      \hline
      Range        & 30 ft \\
      \hline
      Components   & V, S, M \\
      \hline
      Duration     & Concentration, up to 1 minute \\
      \hline
      Area         & Target \\
      \hline
   \end{tabular}
\end{table}

\paragraph{}
A creature of your choice that you can see within range perceives everything as hilariously funny and falls into fits of laughter if this spell affects it. The target must succeed on a wisdom saving throw or fall prone, becoming incapacitated and unable to stand up for the duration. A creature with an Intelligence score of 4 or less isn’t affected.

\paragraph{}
At the end of each of its turns, and each time it takes damage, the target can make another wisdom saving throw. The target had advantage on the saving throw if it’s triggered by damage. On a success, the spell ends.

\newpage

\section*{Invisibility}
\begin{table}[h]
   \centering
   \begin{tabular}{|l|r|}
      \hline
      Level        & II \\
      \hline
      School       & Illusion \\
      \hline
      Casting Time & Action \\
      \hline
      Range        & Touch \\
      \hline
      Components   & V, S, M \\
      \hline
      Duration     & Concentration, up to 1 hour \\
      \hline
      Area         & Target \\
      \hline
   \end{tabular}
\end{table}

\paragraph{}
A creature you touch becomes invisible until the spell ends. Anything the target is wearing or carrying is invisible as long as it is on the target's person. The spell ends for a target that attacks or casts a spell.

\paragraph{At higher levels:}
When you cast this spell using a spell slot of 3rd level or higher, you can target one additional creature for each slot level above 2nd.

\newpage

\section*{Phantasmal Force}
\begin{table}[h]
   \centering
   \begin{tabular}{|l|r|}
      \hline
      Level        & II \\
      \hline
      School       & Illusion \\
      \hline
      Casting Time & Action \\
      \hline
      Range        & 60 ft \\
      \hline
      Components   & V, S, M \\
      \hline
      Duration     & Concentration, up to 1 minute \\
      \hline
      Area         & Target \\
      \hline
   \end{tabular}
\end{table}

\footnotesize{}
\paragraph{}
You craft an illusion that takes root in the mind of a creature that you can see within range. The target must make an intelligence saving throw. On a failed save, you create a phantasmal object, creature, or other visible phenomenon of your choice that is no larger than a 10-foot cube and that is perceivable only to the target for the duration. This spell has no effect on undead or constructs.

\paragraph{}
The phantasm includes sound, temperature, and other stimuli, also evident only to the creature.The target can use its action to examine the phantasm with an Intelligence (Investigation) check against your spell save DC. If the check succeeds, the target realizes that the phantasm is an illusion, and the spell ends.

\paragraph{}
While a target is affected by the spell, the target treats the phantasm as if it were real. The target rationalizes any illogical outcomes from interacting with the phantasm. For example, a target attempting to walk across a phantasmal bridge that spans a chasm falls once it steps onto the bridge. If the target survives the fall, it still believes that the bridge exists and comes up with some other explanation for its fall—it was pushed, it slipped, or a strong wind might have knocked it off.

\paragraph{}
An affected target is so convinced of the phantasm's reality that it can even take damage from the illusion. A phantasm created to appear as a creature can attack the target. Similarly, a phantasm created to appear as fire, a pool of acid, or lava can burn the target. Each round on your turn, the phantasm can deal 1d6 psychic damage to the target if it is in the phantasm's area or within 5 feet of the phantasm, provided that the illusion is of a creature or hazard that could logically deal damage, such as by attacking. The target perceives the damage as a type appropriate to the illusion.
\normalsize{}

\newpage

\section*{Suggestion}
\begin{table}[h]
   \centering
   \begin{tabular}{|l|r|}
      \hline
      Level        & II \\
      \hline
      School       & Enchantment \\
      \hline
      Casting Time & Action \\
      \hline
      Range        & 30 ft \\
      \hline
      Components   & V, M \\
      \hline
      Duration     & Concentration, up to 8 hours\\
      \hline
      Area         & Target \\
      \hline
   \end{tabular}
\end{table}

\paragraph{}
You suggest a course of activity (limited to a sentence or two) and magically influence a creature you can see within range that can hear and understand you. Creatures that can't be charmed are immune to this effect. The suggestion must be worded in such a manner as to make the course of action sound reasonable. Asking the creature to stab itself, throw itself onto a spear, immolate itself, or do some other obviously harmful act ends the spell.

\paragraph{}
The target must make a Wisdom saving throw. On a failed save, it pursues the course of action you described to the best of its ability. The suggested course of action can continue for the entire duration. If the suggested activity can be completed in a shorter time, the spell ends when the subject finishes what it was asked to do.

\paragraph{}
You can also specify conditions that will trigger a special activity during the duration. For example, you might suggest that a knight give her warhorse to the first beggar she meets. If the condition isn't met before the spell expires, the activity isn't performed.

\paragraph{}
If you or any of your companions damage the target, the spell ends.

\newpage

\section*{Hypnotic Pattern}
\begin{table}[h]
   \centering
   \begin{tabular}{|l|r|}
      \hline
      Level        & III \\
      \hline
      School       & Illusion\\
      \hline
      Casting Time & Action \\
      \hline
      Range        & 120 ft \\
      \hline
      Components   & S, M \\
      \hline
      Duration     & Concentration, up to 1 minute\\
      \hline
      Area         & Cube, 30 ft\\
      \hline
   \end{tabular}
\end{table}

\paragraph{}
You create a twisting pattern of colors that weaves through the air inside a 30-foot cube within range. The pattern appears for a moment and vanishes. Each creature in the area who sees the pattern must make a Wisdom saving throw. On a failed save, the creature becomes charmed for the duration. While charmed by this spell, the creature is incapacitated and has a speed of 0.

\paragraph{}
The spell ends for an affected creature if it takes any damage or if someone else uses an action to shake the creature out of its stupor.

\newpage

\section*{Green-Flame Blade}
\begin{table}[h]
   \centering
   \begin{tabular}{|l|r|}
      \hline
      Level        & Cantrip \\
      \hline
      School       & Evocation \\
      \hline
      Casting Time & Action \\
      \hline
      Range        & 5 ft \\
      \hline
      Components   & V, M (a weapon) \\
      \hline
      Duration     & Instantaneous \\
      \hline
      Area         & Target \\
      \hline
   \end{tabular}
\end{table}

\paragraph{}
As part of the action used to cast this spell, you must make a melee attack with a weapon against one creature within the spell's range, otherwise the spell fails. On a hit, the target suffers the attack's normal effects, and green fire leaps from the target to a different creature of your choice that you can see within 5 feet of it. The second creature takes fire damage equal to your spellcasting ability modifier. This spell's damage increases when you reach higher levels.

\paragraph{}
At 5th level, the melee attack deals an extra 1d8 fire damage to the target, and the fire damage to the second creature increases to 1d8 + your spellcasting ability modifier. Both damage rolls increase by 1d8 at 11th level and 17th level.

\newpage

\section*{Enemies abound}
\begin{table}[h]
   \centering
   \begin{tabular}{|l|r|}
      \hline
      Level        & III \\
      \hline
      School       & Enchantment \\
      \hline
      Casting Time & Action \\
      \hline
      Range        & 120 ft \\
      \hline
      Components   & V, S \\
      \hline
      Duration     & Concentration, up to 1 minute\\
      \hline
      Area         & Target \\
      \hline
   \end{tabular}
\end{table}

\paragraph{}
You reach into the mind of one creature you can see and force it to make an Intelligence saving throw. A creature automatically succeeds if it is immune to being frightened. On a failed save, the target loses the ability to distinguish friend from foe, regarding all creatures it can see as enemies until the spell ends. Each time the target takes damage, it can repeat the saving throw, ending the effect on itself on a success. Whenever the affected creature chooses another creature as a target, it must choose the target at random from among the creatures it can see within range of the attack, spell, or other ability it’s using. If an enemy provokes an opportunity attack from the affected creature, the creature must make that attack if it is able to.

\newpage

\section*{Chill Touch}
\begin{table}[h]
   \centering
   \begin{tabular}{|l|r|}
      \hline
      Level        & Cantrip \\
      \hline
      School       & Necromancy \\
      \hline
      Casting Time & Action \\
      \hline
      Range        & 120 ft \\
      \hline
      Components   & V, S \\
      \hline
      Duration     & Instantaneous \\
      \hline
      Area         & Target \\
      \hline
   \end{tabular}
\end{table}

\paragraph{}
You create a ghostly, skeletal hand in the space of a creature within range. Make a ranged spell attack against the creature to assail it with the chill of the grave. On a hit, the target takes 1d8 necrotic damage, and it can't regain hit points until the start of your next turn. Until then, the hand clings to the target.

\paragraph{}
If you hit an undead target, it also has disadvantage on attack rolls against you until the end of your next turn.

\paragraph{}
This spell's damage increases by 1d8 when you reach 5th level (2d8), 11th level (3d8), and 17th level (4d8).

\newpage

\section*{Armor of Agathys}
\begin{table}[h]
   \centering
   \begin{tabular}{|l|r|}
      \hline
      Level        & I \\
      \hline
      School       & Abjuration \\
      \hline
      Casting Time & Action \\
      \hline
      Range        & Self \\
      \hline
      Components   & V, S, M \\
      \hline
      Duration     & 1 hour \\
      \hline
      Area         & Self \\
      \hline
   \end{tabular}
\end{table}

\paragraph{}
A protective magical force surrounds you, manifesting as a spectral frost that covers you and your gear. You gain 5 temporary hit points for the duration. If a creature hits you with a melee attack while you have these hit points, the creature takes 5 cold damage.

\paragraph{At higher levels:}
When you cast this spell using a spell slot of 2nd level or higher, both the temporary hit points and the cold damage increase by 5 for each slot level above 1st.

\newpage

\section*{Shield}
\begin{table}[h]
   \centering
   \begin{tabular}{|l|r|}
      \hline
      Level        & I \\
      \hline
      School       & Abjuration \\
      \hline
      Casting Time & Reaction \\
      \hline
      Range        & Self \\
      \hline
      Components   & V, S \\
      \hline
      Duration     & 1 round \\
      \hline
      Area         & Self \\
      \hline
   \end{tabular}
\end{table}

\paragraph{}
An invisible barrier of magical force appears and protects you. Until the start of your next turn, you have a +5 bonus to AC, including against the triggering attack, and you take no damage from magic missile.

\newpage

\section*{Counterspell}
\begin{table}[h]
   \centering
   \begin{tabular}{|l|r|}
      \hline
      Level        & III \\
      \hline
      School       & Abjuration \\
      \hline
      Casting Time & Reaction \\
      \hline
      Range        & 60 ft \\
      \hline
      Components   & S \\
      \hline
      Duration     & Instantaneous \\
      \hline
      Area         & Target \\
      \hline
   \end{tabular}
\end{table}

\paragraph{}
You attempt to interrupt a creature in the process of casting a spell. If the creature is casting a spell of 3rd level or lower, its spell fails and has no effect. If it is casting a spell of 4th level or higher, make an ability check using your spellcasting ability. The DC equals 10 + the spell's level. On a success, the creature's spell fails and has no effect.

\paragraph{At higher levels:}
When you cast this spell using a spell slot of 4th level or higher, the interrupted spell has no effect if its level is less than or equal to the level of the spell slot you used.

\newpage

\section*{Slow}
\begin{table}[h]
   \centering
   \begin{tabular}{|l|r|}
      \hline
      Level        & III \\
      \hline
      School       & Transmutation \\
      \hline
      Casting Time & Action \\
      \hline
      Range        & 120 ft \\
      \hline
      Components   & V, S, M \\
      \hline
      Duration     & Concentration, up to 1 minute\\
      \hline
      Area         & Cube, 40 ft, up to 6 targets\\
      \hline
   \end{tabular}
\end{table}

\paragraph{}
You alter time around up to six creatures of your choice in a 40-foot cube within range. Each target must succeed on a Wisdom saving throw or be affected by this spell for the duration.

\paragraph{}
An affected target's speed is halved, it takes a -2 penalty to AC and Dexterity saving throws, and it can't use reactions. On its turn, it can use either an action or a bonus action, not both. Regardless of the creature's abilities or magic items, it can't make more than one melee or ranged attack during its turn.

\paragraph{}
If the creature attempts to cast a spell with a casting time of 1 action, roll a d20. On an 11 or higher, the spell doesn't take effect until the creature's next turn, and the creature must use its action on that turn to complete the spell. If it can't, the spell is wasted.

\paragraph{}
A creature affected by this spell makes another Wisdom saving throw at the end of its turn. On a successful save, the effect ends for it.


\end{document}
